---
title: "Khyati Naik: Data 605 - HW13"
date: "Nov 25, 2023"
---

# 1. Use integration by substitution to solve the integral below

\[ \int 4e^{-7x} \,dx \]


**Step 1: Choose a Substitution**

Let \( u = -7x \), then \( du = -7 \,dx \).

**Step 2: Substitute and Simplify**

\[ \int 4e^{-7x} \,dx = \int 4e^u \left(-\frac{1}{7}\right) \,du \]

**Step 3: Integrate and Back-Substitute**

\[ = -\frac{4}{7} \int e^u \,du \]

Now, integrate \( \int e^u \,du \) and back-substitute. The result obtained after integration is:

\[ -\frac{4}{7} e^{-7x} + C \]

where \(C\) is the constant of integration.


So, the solution to the integral \( \int 4e^{-7x} \,dx \) is \( -\frac{4}{7} e^{-7x} + C \), where \(C\) is the constant of integration.


# 2. Biologists are treating a pond contaminated with bacteria. The level of contamination is changing at a rate of dN/dt = (-3150/t^4) - 220dt  bacteria per cubic centimeter per day, where t is the number of days since treatment began. Find a function N( t ) to estimate the level of contamination if the level after 1 day was 6530 bacteria per cubic centimeter.


The level of contamination is changing at a rate of:

\[ \frac{dN}{dt} = -\frac{3150}{t^4} - 220 \]

**Integrate the Equation**

Assuming \(N\) is a function of \(t\) (\(N(t)\)), integrate the equation:

\[ \int \frac{1}{N} \, dN = \int \left(-\frac{3150}{t^4} - 220\right) \, dt \]

\[ N(t) = {\frac{1050}{t^3} - 220t + C} \]

When t=1 and N(t)=6530
N(1) = 1050 -220 +C=6530
Therefore, C=5700

\[ N(t) = {\frac{1050}{t^3} - 220t + 5700} \]


# 3. Find the total area of the red rectangles in the figure below, where the equation of the line is f(x) = 2x - 9.

```{r}
# Define the function
function_integ <- function(x) {
    2 * x - 9
}

# Integrate the function from x = 4.5 to x = 8.5
area <- integrate(function_integ, 4.5, 8.5)

# Display the total area
area
```
The total area of the red rectangles in the figure is 16.

# 4. Find the area of the region bounded by the graphs of the given equations.
y = x^2 - 2x - 2, y = x + 2

Solution:

x^2 - 2x - 2=x+2
x2−3x−4=0
(x−4)(x+1)=0
x=4,x=−1

Subtract the two equations.
x+2−(x^2−2x−2)
x+2−x^2+2x+2
−x^2+3x+4
Integrate this equation using intersection points as interval.

```{r}
function_integ <- function(x)
{
  -x^2 + 3*x + 4
}
integrate(function_integ, lower = -1, upper = 4)
```
Thus, the area of the region bounded by the graphs of the given equations is 20.833.

# 5. A beauty supply store expects to sell 110 flat irons during the next year. It costs $3.75 to store one flat iron for one year. There is a fixed cost of $8.25 for each order. Find the lot size and the number of orders per year that will minimize inventory costs.

```{r}
# Given values
D <- 110  # Demand per year
S <- 8.25  # Setup cost per order
H <- 3.75  # Holding cost per unit per year

# Economic Order Quantity (EOQ) formula
Q_star <- sqrt((2 * D * S) / H)

# Number of orders per year
Num_Orders <- D / Q_star

Q_star
Num_Orders
```
The lot size of 22 and the number of orders per year of 5 will minimize inventory costs.

# 6. Use integration by parts to solve the integral below.

\[ \int \ln(9x) \cdot x^6 \,dx \]

Let:
\[ u = \ln(9x) \]
\[ dv = x^6 \,dx \]

Now, calculate \( du \) and \( v \):
\[ du = \frac{1}{\ln(9x)} \cdot \frac{d}{dx} (9x) \,dx = \frac{\ln(9x)}{9} \cdot 9 \,dx = \ln(9x) \,dx \]
\[ v = \frac{1}{7}x^7 \]

Now, apply the integration by parts formula:
\[ \int \ln(9x) \cdot x^6 \,dx = uv - \int v \,du \]

\[ = \ln(9x) \cdot \frac{1}{7}x^7 - \int \frac{1}{x}\frac{1}{7}x^7 dx \]

\[ = \ln(9x) \cdot \frac{1}{7}x^7 - \frac{1}{49}x^7 + C \]


# 7. Determine whether f(x) is a probability density function on the interval [1,e^6]. If not, determine the value of the definite integral.
f(x) = 1/6x

**Determining if f(x) is a Probability Density Function**

To check if \( f(x) = \frac{1}{6x} \) is a probability density function (pdf) on the interval \([1, e^6]\), we need to verify two conditions:

1. Non-negativity: \( f(x) \) is non-negative for all \( x \) in the interval.
2. Normalization: The integral of \( f(x) \) over the entire range is equal to 1.

**Non-negativity**

For \( f(x) = \frac{1}{6x} \), \( f(x) \) is non-negative as long as \( x > 0 \). In the given interval \([1, e^6]\), \( x \) is always greater than 0, so this condition is satisfied.

**Normalization**

We need to check whether the integral of \( f(x) \) over the interval \([1, e^6]\) is equal to 1.

\[
\int_{1}^{e^6} \frac{1}{6x} \,dx = \frac{1}{6} \int_{1}^{e^6} \frac{1}{x} \,dx
\]

The integral \( \int \frac{1}{x} \,dx \) is equal to \( \ln(x) \), so:

\[
\frac{1}{6} \int_{1}^{e^6} \frac{1}{x} \,dx = \frac{1}{6} [\ln(x)]_{1}^{e^6}
\]

Evaluate this:

\[
\frac{1}{6} [\ln(e^6) - \ln(1)] = \frac{1}{6} [6 - 0] = 1
\]

So, the integral is equal to 1, and the second condition is satisfied.

Therefore, \( f(x) = \frac{1}{6x} \) is a valid probability density function on the interval \([1, e^6]\).

